\section{Introduction}

Why is it different to deploy sensors in an urban environment than deploying them on a rural setting. The main difference lies on the presence of people and their right to privacy. Secondary differences lie on the easier access to power, increased chance of human activity interference to communication (in case it is wireless).

Sensors can measure a variety on metrics. Some are not consider privacy intrusive such as temperature, humidity or wind speed. Others can be considered intrusive based on their usage, post processing, in combination with other metrics or aggregation from different locations. Such potentially intrusive sensing data are capture of image, sound, video, wireless traffic or rfid card presence.

Based on the above, a list of design considerations can be constructed.


\paragraph{First order design considerations}

\begin{itemize}
 \item Privacy: Protect the privacy of people that come in sensing vicinity from the embedded devices
 \item Security: Preserve the integrity of the system, the data and assure no privacy breach
 \item Survivability: Survive hardware or software failures due to bugs, weather conditions, or natural wear-out.
 \item Power Efficiency: Minimize amount of required operational power, stay inside provided energy budget, and instrument power source (battery, renewable energy) to increase availability.
\end{itemize}

\paragraph{Second order design considerations}

\begin{itemize}
 \item Network Communication Efficiency: Minimize network traffic 
 \item Cost: Hardware, software, and maintenance cost
\end{itemize}

Following we will investigate different design choices that try to improve the above goals in the best possible way. The assumed scheme would be WAggLE (see below) and the deployment location the city of Chicago.

\subsection{The WAggLE architecture}

\subsubsection{Cloud}

From the Node Controller side it is where data are send and commands are received from.
From the external user side it is where commands are send to so that they are forwarded to the nodes and also the public access point of sensor data
From the Guest Node side the cloud is unreachable

\subsubsection{Node Controller}

From the external user side it is not accessible
From the cloud side it is the source of sensor data and the destination for commands
From the guest node side it is the destination of collected data and a liaison for requesting data of other sensors


\subsubsection{Guest Node}

From the cloud and external use side the Guest Node is unreachable
From the Node Controller side it is the source of sensor data and destination of requested data.


\subsection{City of Chicago}

\url{https://en.wikipedia.org/wiki/Climate_of_Chicago}

Low temperatures on most of the year.

Can get quite hot during summer.

High humidity.

Rainfall.

Snow.

Stray Bullets.
