\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[11pt,letter]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[colorlinks=false, pdfborder={0 0 0}]{hyperref}
\usepackage{cleveref}
\usepackage{bytefield}
\usepackage{pdflscape}
\usepackage[ margin=1in]{geometry}
\usepackage{framed}
\usepackage[table]{xcolor}
\newcommand{\guestnode}{hosted\ node\ }
\newcommand{\zerolenmessage}{2--3 & Length of Message Body & $0x00$ $0x00$ & Empty Message. \\ \hline}
\usepackage{fancyhdr}
\usepackage[yyyymmdd,hhmmss]{datetime}
\usepackage{xcolor}
\usepackage{sectsty}

\clubpenalty=10000
\widowpenalty=10000

\definecolor{rahmen}{RGB}{0,73,114}
\definecolor{grund}{RGB}{238,241,251}
\definecolor{schrift}{RGB}{116,45,19}
\sectionfont{\color{black}}
\subsectionfont{\color{rahmen}}
\subsubsectionfont{\color{schrift}}

\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\cfoot{\thepage}
\rfoot{\today\ \currenttime}
\begin{document}
\title{Messaging Protocol Version 0.4}
\author{The Waggle Team}
\date{2014-2016}
\maketitle
\noindent
For all the messages sent using protocol Version 0.4, the following fields will be standard across
all headers, and across all devices.
\begin{center}
    \rowcolors{1}{rahmen!8}{white}
    \begin{tabular}{ | l | l | p{3cm} | p{5cm} |}
    \hline
    \hline
    \textbf{Byte Field} & \textbf{Field Name} & \textbf{Value} & \textbf{Notes} \\ \hline \hline
    0 & Protocol Version & $0x04$ & The Current Version of Protocol is Major 0, Minor 4. \\    \hline
    10 & Extended Header & $0x01$ & Set if \textit{Optional Key} is presented. \\    \hline
    11 & Optional Keys & $0xzz$ & 8--bit flags. \\    \hline
    \end{tabular}
\end{center}
\begin{framed}
\textbf{Notes:}
\\
\begin{itemize}
\item This version of protocol provides data transmission between beehive server and nodes. This supports three communication scenarios: 1) data transmission, 
2) request/response, and 3) potentially a request initiated from the beehive server.
\item The major differences in this version are --
\begin{itemize}
\item The size of message header is 40 Bytes including CRC16 and is reduced by resizing the sender and responder's unique IDs by 2 bytes each.
\item A plugin in both server and node sides can be revealed by looking at the plugin ID in message.
\item Additional message types such as response of a request.
\item Data are serialized and compressed in message body.
\item Sending a large message exceeding 1 kB is supported using the $Optional\_key$ field.
\end{itemize}
\end{itemize}
\end{framed}

\input{Packet_Format}
\input{Sensor_Queries}
\input{Message_Types}



\end{document}
